---
title: "Parallel processing on windows"
author: "Jesse Tweedle"
date: 2017-12-17T15:33:11-05:00
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

<!-- Doesn't need a title -->

<!-- ## Closure hack for windows parallel processing -->

I use the term hack bc it sounds cool. Closure rly legit programming technique to solve this prob.  

I need to write some real shit here. Maybe a windows related post? Using closures to skip the clusterExport step? Functions?

Doesn't fork, starts new. Can use mclapply.hack thing, but that exports everything, takes super long time if you have lots of data in global env. Want to pass only what we need. Including libraries, functions.

<!-- You can embed an R code chunk like this: 


```{r cars}
summary(cars)
fit <- lm(dist ~ speed, data = cars)
fit
```

# Including Plots

You can also embed plots. See Figure \@ref(fig:pie) for example:

```{r pie, fig.cap='A fancy pie chart.', tidy=FALSE}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid', "eighty charact"),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```
-->