---
title: 'Jalen v. Shaq as baby names'
author: "Jesse Tweedle"
date: '2018-01-27'
slug: jalen-v-shaq
categories: ["r"]
tags: ["r", "nba", "babynames"]
description: 'Was Jalen Rose really the first Jalen?'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, fig.height = 4, fig.width = 8)
library(tidyverse)
library(ggplot2)
```

Was [Jalen Rose](https://en.wikipedia.org/wiki/Jalen_Rose) really the first Jalen? He claims his mother was the first to make up the name, a combo of his father's (the [NBA player Jimmie Walker](https://en.wikipedia.org/wiki/Jimmy_Walker_(basketball,_born_1944))) and uncle's (Leonard) names. And now we see them everywhere: [Jalen Hurts](https://en.wikipedia.org/wiki/Jalen_Hurts) (the Alabama QB that was benched in the national title game a few weeks ago) and [Jalen Ramsey](https://en.wikipedia.org/wiki/Jalen_Ramsey) (Jags cornerback). Was he really? On the other hand, Shaq is an equally distinct name, what happened after he got famous?

This project is almost one line. Get a dataset of babyname counts (via [Hadley's `babynames`](https://github.com/hadley/babynames)), filter for Jalen and Shaq(uille), and plot them over time. But first we'll see if there were any Jalens before he started at Michigan in 1991:

``` {r}
library(babynames)
babynames %>% 
  filter(name == "Jalen") %>% 
  group_by(name, year) %>% 
  summarize(n = sum(n))
```

Hmm. Jalen was born in 1973 but it doesn't show up until 1976. I'm guessing the `babynames` dataset is censored at 5 to preserve confidentiality or something. So he might've been the first! But unless his mother was promoting the name when he was 3, others invented 'Jalen' simultaneously. 

But you can definitely see 'Jalen' uptick when he hit the national stage. What does the trend look like? How does it compare to a similarly famous and unique name?

``` {r}
babynames %>% 
  filter(name %in% c("Jalen", "Shaquille")) %>% 
  group_by(name, year) %>% summarize(n = sum(n)) %>% 
    ggplot(aes(x = year, y = n, group = name, colour = name)) + 
    geom_point() + geom_line() +
    labs(title = "Jalen v Shaq as baby names", xlab = "year", ylab = "Number of babies") + 
    guides(fill=guide_legend(title="Name"))
```

Daaaaaaaang Jalen `r emo::ji('100')`. #gtpwtw since 1991. Around 1992-3 Shaq looks just as popular, then ppl realize it's crazy to call your baby Shaquille, he's too famous! E.g., the only person who can name a baby Lebron is Lebron. (Err, wait! There were 6 Lebrons born in 1925?!)

But Jalen sticks around and hits a maximum of `r babynames %>% filter(name == "Jalen") %>% group_by(name, year) %>% summarize(n = sum(n)) %>% pull(n) %>% max()` per year in 2000. He's slid off since then, but still, very impressive Jalen. 

#1 in my babyname directory, #1 in my heart.